Learning a functional distance between organizations with unsupervised embeddings of scientific mobility

### 2020-07-31 paper/current/main.tex
remainder of Sadamori's comments

Just as the embedding space makes it possible to ``zoom in'' on subsets of organizations, it is also possible to ``zoom out'' by aggregating organizational vectors.
We can examine the country-level structure that governs scientific mobility.
To this end, we define the representative vector of each country as the average of their organizational vectors.
We then perform hierarchical clustering of nations that have at least 25 organizations represented in the embedding space based on their cosine similarity, excluding the United States which is a dominant hub well connected to most other countries (see Fig. \ref{fig:clustering}a).
The six identified clusters roughly correspond to countries in East-Asia (orange), Scandinavia (dark purple), the British Commonwealth (light purple), and Central and Eastern Europe (light blue), and then two remaining diverse clusters contain a mix of European, Latin American, and Mediterranean countries.
These clusters demonstrate the importance of not only geography but also linguistic ties between countries.

@@

%
% Figure - CLUSTERING
%
\begin{figure}[p!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Clustering/cluster_country.pdf}
	\caption{
		\textbf{Geography, then language, conditions international mobility.}
		\textbf{a.} Hierarchically clustered similarity matrix of country vectors, aggregated as the mean of all organization vectors within the country that has more than 25 organizations, excluding the United States.
		Color of matrix cells corresponds to the cosine similarity between country vectors.
		Color of country names corresponds to their cluster.
		Color of three cell rows separated from the matrix corresponds to, from left to right, the region of the country, the language family~\autocite{ethnologue}, and the individual language.
		\textbf{b.}	Element-centric cluster similarity~\autocite{gates2019element} reveals the factors dictating hierarchical clustering.
		Region better explains the grouping of country vectors at higher levels of the clustering.
		Language family, and then the most widely-spoken language, better explain the grouping of countries at more fine-grained level.
	}
	\label{fig:clustering}
\end{figure}

@@

Prestige hierarchy is known to underpin the dynamics of scientific mobility~\autocite{deville2014career, clauset2015hierarchy}, but can the hierarchy be encoded in the embedding space, to which no explicit information on prestige is given and for which the direction of mobility is not known?
We test this by exploiting the geometric properties of the embedding space with SemAxis~\autocite{an2018semaxis}, an approach originally designed to characterize the semantics of words in word embeddings.
Here, we use SemAxis to operationalize the abstract notion of academic prestige, defining each pole of the axis using high- and low-ranked universities.
To define prestige of ranked U.S. ranked univeristies, we use the Times Ranking of World Universities (we also use research impact from the Leiden Rankings~\autocite{waltman2012leidenrankings}, see \SI);
the high-rank pole is defined as the mean vector of the top five U.S. universities according to the rankings, whereas the low-ranked pole is defined using the five bottom-ranked (geographically-matched by U.S. census region) universities.

The embedding space encodes the prestige hierarchy of U.S. universities well enough to re-construct university rankings. 
We derive a ranking for universities using the SemAxis prestige axis and compare them to traditional university rankings.
The embedding-based ranking is defined from the position of organizations along the prestige axis, with the top-ranked being the nearest to the high-ranked pole.
The embedding-based ranking is strongly correlated with the Times ranking (Spearman's $\rho = 0.73$, Fig.~\ref{fig:semaxis}a).
For reference, the embedding-based ranking achieves correlation with the Times Ranking comparable to that between the Times ranking and the publication impact scores from the Leiden Ranking~\autocite{waltman2012leidenrankings}, a well-known bibliometrically-informed university ranking (Spearman's $\rho = 0.87$, Fig.~\ref{fig:semaxis}b).
The correlation between the embedding-based ranking and the Times ranking is robust regardless of the number of organizations used to define the axes (Fig. SI~\ref{fig:supp:semaxis_compare}), such that even using only the single top-ranked and bottom-ranked universities produces a ranking that is significantly correlated with the Times Ranking (Spearman's $\rho = 0.46$, Fig.~\ref{fig:supp:semaxis_compare}a).
Performance is also comparable to network-driven measures such as degree strength (Spearman's $\rho = 0.73$) and eigenvector centrality (Spearman's $\rho = 0.76$, see \SI). 
The strongest outliers that were ranked more highly in the Times Ranking than in the embedding-based ranking tend to be large state universities such as Arizona State University and  University of Florida.
Those ranked higher in the embedding based ranking tend to be relatively-small universities near major urban areas such as the University of San Francisco and University of Maryland Baltimore County.
This suggests that the organization embedding captures the latent prestige hierarchy inherent to academic mobility with a high degree of resolution and reliability, even without providing any explicit information on prestige~\autocite{deville2014career, clauset2015hierarchy}.

The axes can further be visualized to examine the relative position of organizations along the prestige axis, and along a geographic axis between California and Massachusetts that aids visualization.
Prestigious universities such as Columbia, Stanford, MIT, Harvard, and Rockefeller are positioned towards the top of the axis (Fig.~\ref{fig:semaxis}c).
Universities at the bottom of this axis tend to be regional universities with lower national profiles (yet still ranked by Times Higher Education) and with more emphasis on teaching, such as Barry University and California State University at Long Beach.
By projecting other types of organizations onto the prestige axis, SemAxis offers a new means of reasoning about the prestige of organizations for which rankings are often low-resolution or incomplete, such as regional and liberal arts universities, seminaries, musical schools, and community colleges (Fig.~\ref{fig:semaxis}d), or for which no such rankings exist, such as government organizations (Fig.~\ref{fig:semaxis}d), and research institutes (Fig.~\ref{fig:semaxis}e).

%
% Figure - SemAxis results
%
\begin{figure}[hp!]
	\centering
	\includegraphics[width=0.7\textwidth]{\figdir/SemAxis/semaxis_fig.pdf}
	\caption{
		\textbf{Embedding captures latent geography and prestige hierarchy.}
		\textbf{a.} Comparison between the ranking of organizations by their rank in the Times ranking and the embedding rank derived using SemAxis with the poles defined using the top five, and geographically-matched bottom five universities ranked by the 2018 Times Higher Education ranking of U.S. Universities .
		Un-filled points are those top and bottom five universities used to define the poles of the axis.
	  	Even when considering only a total of ten organization vectors, the estimate of the Spearman's rank correlation between the embedding and Times ranking is $\rho = 0.73$ (n = 145), and it increases if we include more top-and-bottom ranked universities (Fig.~\ref{fig:supp:semaxis_compare}).
	  	\textbf{b.} Times ranking is correlated with Leiden Ranking of U.S. universities with Spearman's $\rho = 0.87$
		\textbf{c-f.} Illustration of SemAxis projection along two axes;  \textit{latent geographical axis}, from California to Massachusetts (left to right) and the \textit{prestige axis}.
		Shown for U.S. Universities (\textbf{c}), regional and liberal arts colleges (\textbf{d}), Research institutes (\textbf{e}), and Government organizations (\textbf{f}).
		Full organization names listed in Table~\ref{table:supp:orglabels}.
	}
	\label{fig:semaxis}
\end{figure}

We also find that the size (L2 norm) of the organization embedding vectors provides insights into the characteristics of organizations.
Up to a point (around 1,000 researchers), the size of U.S. organization's vectors tends to increase proportional to the number of researchers (both mobile and non-mobile) within that organization with published work;
these organizations are primarily teaching-focused institutions, agencies, and hospitals that either are not ranked or have a low ranking.
However, at around 1,000 researchers, the size of the vector decreases as the number of researchers increases.
These organizations are primarily research-intensive and prestigious universities with higher rankings, research outputs, R\&D funding, and doctoral students (Fig.~\ref{fig:supp:length_vs_metaInfo}).
A similar pattern has been observed in applications of neural embedding to natural language, wherein the size of word vectors was found to represent the word ``significance''~\autocite{schakel2015measuring}.
If a word appears frequently in a specific context, the norm of the word vector is likely to be large, because a large norm is associated with how predictably the word occurs given a context.
However, if the word in question is universal, appearing frequently in many different contexts, it cannot have a large norm as it lacks a strong association with a particular context.
Likewise, an organization such as Harvard with a small norm appears in many contexts, alongside many different organizations in affiliation trajectories---it is well-connected.
The concavity emerges in part from the relationship between the size of the vector and the expected connectedness of the organization given its size (Fig.~\ref{fig:length}a, inset, $R^{2} = 0.17$).
Large, prestigious, and well-funded research universities such as Princeton and Harvard have smaller vector sizes because they appear in many different contexts compared to more teaching-focused organizations such as NY Medical College, and University of Michigan at Flint.
Some universities, such as the University of Alaska at Fairbanks, have surprisingly small vectors, perhaps as a result of its remote location and unique circumstances.

This same convex-curve repeats across many countries such as China, which closely mirrors that of the United States (Fig.~\ref{fig:length}b).
Smaller but scientifically advanced countries such as Australia, and other large populous countries such as Brazil exhibit curves similar to the United States (Fig.~\ref{fig:length}b, inset), though their exact position and size varies.
Other nations exhibit different curves, such as Egypt which has an``S''-shaped curve---meaning that their largest organizations are not as well connected as other countries' largest organizations. 
Similar patterns can be found across many of the 30 of the countries with the most total researchers (Fig.~\ref{fig:concave30}), though we can only speculate on what their variations mean (see \SI).
Neural embeddings encode a relationship between organization size and vector length that is universal across countries but with potentially meaningful nuances, helping the embedding to capture the landscape of global scientific mobility.

%
% Figure - Boomerang
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Math/boomerang.pdf}
	\caption{
		\textbf{Size of organization embedding vectors captures prestige and size of organizations.}
		\textbf{a.} Size (L2 norm) of organization embedding vectors compared to the number of researchers for U.S. universities.
		Color indicates the rank of the university from the Times ranking, with 1 being the highest ranked university.
		Uncolored points are universities not listed on a Times ranking.
		A convex-shape emerges, wherein larger universities tend to be more distant from the origin (large L2 norm); however, the more prestigious the university, the closer it is to the origin (small L2 norm).
		The L2 norm is correlated weakly with the ratio of organizations' number of connections (non-zero flows) to the number of researchers ($R^{2} = 0.17$, \textbf{inset}).
		The red line is the line of the best fit, and 99\% confidence intervals are shown, though are too small to be visible.
		\textbf{b.} This convex pattern repeats globally for all organizations across all countries, shown for the United States, China, and others,
		The loess regression lines are shown for each selected country, and for the aggregate of remaining countries, with 99\% confidence intervals.
		Loess lines are also shown for organizations in Australia, Brazil, and Egypt (\textbf{inset}).
	}
	\label{fig:length}
\end{figure}

@@

Neural embeddings offer a novel, data-driven solution for learning an effective and robust functional distance between places based on trajectory data, encoding the complex and multi-faceted nature of mobility.
We demonstrated how the embedding space encodes various aspects of scientific mobility, such as the global and regional geography, shared languages, and the hierarchy of organizational prestige, even without explicit information on these factors.
While scientific mobility is used as a case study, this approach may be broadly applicable to other kinds of mobility data, such as general human migration, transit-network mobility, and more.
Moreover, this approach can be used to learn a functional distance between entities for which no physical analog exists, such as between occupational categories based on individual's career trajectories.
In addition to providing a functional distance that supports modeling and predicting mobility patterns, the structure of the neural embedding space is amenable to a range of unique applications for studying mobility.
As we have shown, the embedding space allows the visualization of the complex structure of scientific mobility at high resolution across multiple scales, providing the largest and most detailed map of global scientific mobility to date.
Other techniques, such as SemAxis~\autocite{an2018semaxis}, also allow us to quantify abstract notions in mobility such as academic prestige, and can also be generalized to other abstract axes.
Investigation of the structure of the embedding space, such as the vector size, reveals unique and universal patterns based on organization's size and connectedness and could contribute to deeper understandings of scientific mobility.
Neural embedding opens the door for new approaches to represent and study scientific mobility, and also mobility more generally.

In spite of its promise, this approach and analysis has several limitations.
First, the current \textit{skip-gram word2vec} model can not encode directional information, meaning that embedding will be less effective at capturing mobility for which directionality is critical.
Second, the neural embedding approach is most useful in cases of mobility between discrete geographic units such as between countries, cities, and businesses; this approach is less useful in the case of mobility between locations represented using geographic coordinates, such as in the modeling of animal movements.
Neither of these methodological limitations is insurmountable, and future work can aim to incorporate directionality and identify meaningful representations of continuous mobility data.
Finally, the case of scientific mobility presents domain-specific limitations.
Reliance on bibliometric metadata means that we capture only long-term mobility such as migration, rather than the array or more frequent short-term mobility such as conference travel and temporary visits.
The kinds of mobility we do capture—migration and co-affiliation, are treated identically by our model even though they are qualitatively different.
Also, our data might further suffer from survivorship bias: researchers at prestigious organizations tend to have more publications, leading to these organizations appearing more frequently in affiliation trajectories.

Mobility and migration are at the core of human nature and history~\autocite{shah2020migration}, driving societal phenomena as diverse as epidemics\autocite{kraemer2020covid, truscott2012epidemicgravity, xia2005measlesgravity} and science and innovation~\autocite{kaiser2018innovation, armano2017innovation, sugimoto2017mostimpact, petersen2018multiscale, franzoni2014advantage, morgan2018prestige, rodrigues2016mobility}.
However, the paradigm of scientific migration may be changing.
Traditional hubs of migration have experienced many politially-motivated policy changes that affect scientific mobility, such as travel restrictions in the U.S. and U.K.~\autocite{chinchilla2018travelban}.
At the same time, other nations, such as China, are growing into major scientific powers and attractors of talent~\autocite{tollefson2018china, cao2020returning}.
Unprecedented health crises such as COVID-19 pandemic threaten to bring drastic global changes to travel and migration by tightening borders and halting travel.
With the changing paradigm of global mobility we need now, more than ever, new tools and approaches to capture and understand human mobility in order to inform sensible, effective, sustainable, and humane policies.

@@

The performance of the embedding ranking in reconstructing the Times ranking is comparable to that of network-derived measures such as degree strength (Spearman's $\rho = 0.73$, Fig.~\ref{fig:supp:centrality_times_compare}a) and eigencentrality centrality (Spearman's $\rho = 0.76$, Fig.~\ref{fig:supp:centrality_times_compare}b).
However, while both embedding- and network-based measures relate to university prestige, they are qualitatively and quantitatively different.
The embedding-ranking of U.S. universities is less correlated with degree strength (Spearman's $\rho = 0.45$, Fig.~\ref{fig:supp:centrality_semaxis_compare}a) and eigenvector  centrality (Spearman's $\rho = 0.55$) than with the Times Ranking itself (Spearman's $\rho = 0.73$, Fig.~\ref{fig:supp:centrality_semaxis_compare}b). 
The embedding ranking over-ranks large research-intensive universities such as North Carolina State University, University of Florida, and Texas A\&M University, whereas the network-derived ranking over-ranks smaller, more specialized universities such as Brandeis University, Yeshiva University, and University of San Francisco. 
This suggests that the embedding encodes information on prestige hierarchy at least as well as a network representation, but that the information encoded is something different from centrality

>>>>

Just as the embedding space makes it possible to ``zoom in'' on subsets of organizations, it is also possible to ``zoom out'' by aggregating organizational vectors.
We can examine the country-level structure that governs scientific mobility.
To this end, we define the representative vector of each country as the average of their organizational vectors.
We then perform hierarchical clustering of nations that have at least 25 organizations represented in the embedding space based on their cosine similarity, excluding the United States which is a dominant hub well connected to most other countries (see Fig. \ref{fig:clustering}a).
The six identified clusters roughly correspond to countries in East-Asia (orange), Scandinavia (dark purple), the British Commonwealth (light purple), and Central and Eastern Europe (light blue), and then two remaining diverse clusters contain a mix of European, Latin American, and Mediterranean countries.
The cluster structure shows that not only geography but also linguistic ties between countries are related to scientific mobility.

@@

%
% Figure - CLUSTERING
%
\begin{figure}[p!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Clustering/cluster_country.pdf}
	\caption{
		\textbf{Geography, then language, conditions international mobility.}
		\textbf{a.} Hierarchically clustered similarity matrix of country vectors aggregated as the mean of all organization vectors within the country, excluding the United States and countries with less than or equal to 25 organizations.
		Color of matrix cells corresponds to the cosine similarity between country vectors.
		Color of country names corresponds to their cluster.
		Color of three cell rows separated from the matrix corresponds to, from left to right, the region of the country, the language family~\autocite{ethnologue}, and the individual language.
		\textbf{b.}	Element-centric cluster similarity~\autocite{gates2019element} reveals the factors dictating hierarchical clustering.
		Region better explains the grouping of country vectors at higher levels of the clustering.
		Language family, and then the most widely-spoken language, better explain the grouping of countries at more fine-grained level.
	}
	\label{fig:clustering}
\end{figure}

@@

Prestige hierarchy is known to underpin the dynamics of scientific mobility, where researchers tend to move to similar or less prestigious organizations~\autocite{deville2014career, clauset2015hierarchy}. Does the embedding space, to which no prestige information is given, encode a prestige hierarchy?
We test this by exploiting the geometric properties of the embedding space with SemAxis~\autocite{an2018semaxis}.
Here, we use SemAxis to operationalize the abstract notion of academic prestige, defining an axis in the embedding space where poles are defined using known high- and low-ranked universities.
To define prestige of ranked U.S. ranked univeristies, we use the Times Ranking of World Universities (we also use research impact from the Leiden Rankings~\autocite{waltman2012leidenrankings}, see \SI);
the high-rank pole is defined as the mean vector of the top five U.S. universities according to the rankings, whereas the low-ranked pole is defined using the five bottom-ranked (geographically-matched by U.S. census region) universities.
We derive an embedding-based ranking for universities based on their nearness to the high-ranked pole.

The embedding space encodes the prestige hierarchy of U.S. universities that are coherent with real-world university rankings.
The embedding-based ranking is strongly correlated with the Times ranking (Spearman's $\rho = 0.73$, Fig.~\ref{fig:semaxis}a).
For reference, the embedding-based ranking achieves correlation with the Times ranking comparable to that between the Times ranking and the publication impact scores from the Leiden Ranking~\autocite{waltman2012leidenrankings}, a bibliometrically-based university ranking (Spearman's $\rho = 0.87$, Fig.~\ref{fig:semaxis}b).
The correlation between the embedding-based ranking and the Times ranking is robust regardless of the number of organizations used to define the axes (Fig. SI~\ref{fig:supp:semaxis_compare}), such that even using only the single top-ranked and bottom-ranked universities produces a ranking that is significantly correlated with the Times ranking (Spearman's $\rho = 0.46$, Fig.~\ref{fig:supp:semaxis_compare}a).
The correlation is also comparable to network-driven measures such as degree strength (sum of edge weights, Spearman's $\rho = 0.73$) and eigenvector centrality (Spearman's $\rho = 0.76$, see \SI). 
The strongest outliers that were ranked more highly in the Times ranking than in the embedding-based ranking tend to be large state universities such as Arizona State University and  University of Florida.
Those ranked higher in the embedding based ranking tend to be relatively-small universities near major urban areas such as the University of San Francisco and University of Maryland Baltimore County.
This suggests that organizations in the embedding space are arranged according to prestige, even when the erepresentation is learned using data without explicit infomration on the direction of mobility (as in other formal models~\autocite{clauset2015hierarchy}), or prestige.

The axes can further be visualized to examine the relative position of organizations along the prestige axis, and along a geographic axis between California and Massachusetts to aid visualization.
Prestigious universities such as Columbia, Stanford, MIT, Harvard, and Rockefeller are positioned towards the top of the axis (Fig.~\ref{fig:semaxis}c).
Universities at the bottom of this axis tend to be regional universities with lower national profiles (yet still ranked by Times Higher Education) and with more emphasis on teaching, such as Barry University and California State University at Long Beach.
By projecting other types of organizations onto the prestige axis, SemAxis offers a new means of reasoning about the prestige of organizations for which rankings are often low-resolution or incomplete, such as regional and liberal arts universities, seminaries, musical schools, and community colleges (Fig.~\ref{fig:semaxis}d), or for which no such rankings exist, such as government organizations (Fig.~\ref{fig:semaxis}d), and research institutes (Fig.~\ref{fig:semaxis}e).

%
% Figure - SemAxis results
%
\begin{figure}[hp!]
	\centering
	\includegraphics[width=0.7\textwidth]{\figdir/SemAxis/semaxis_fig.pdf}
	\caption{
		\textbf{Embedding captures latent geography and prestige hierarchy.}
		\textbf{a.} Comparison between the ranking of organizations by their rank in the Times ranking and the embedding rank derived using SemAxis pointing from the top five to geographically-matched bottom five universities ranked by the 2018 Times Higher Education ranking of U.S. Universities.
		Un-filled points are those top and bottom five universities used span the axis.
	  	Even when considering only a total of ten organization vectors, the estimate of the Spearman's rank correlation between the embedding and Times ranking is $\rho = 0.73$ ($n = 145$), which increases when more top-and-bottom ranked universities are included (Fig.~\ref{fig:supp:semaxis_compare}).
	  	\textbf{b.} Times ranking is correlated with Leiden Ranking of U.S. universities with Spearman's $\rho = 0.87$
		\textbf{c-f.} Illustration of SemAxis projection along two axes;  \textit{latent geographical axis}, from California to Massachusetts (left to right) and the \textit{prestige axis}.
		Shown for U.S. Universities (\textbf{c}), regional and liberal arts colleges (\textbf{d}), Research institutes (\textbf{e}), and Government organizations (\textbf{f}).
		Full organization names listed in Table~\ref{table:supp:orglabels}.
	}
	\label{fig:semaxis}
\end{figure}

We also find that the size (L2 norm) of the organization embedding vectors provides insights into the characteristics of organizations.
Up to a point (around 1,000 researchers), the size of U.S. organization's vectors tends to increase proportionally to the number of researchers (both mobile and non-mobile) within that organization with published work;
these organizations are primarily teaching-focused institutions, agencies, and hospitals that either are not ranked or have a low ranking.
However, at around 1,000 researchers, the size of the vector decreases as the number of researchers increases.
These organizations are primarily research-intensive and prestigious universities with higher rankings, research outputs, R\&D funding, and doctoral students (Fig.~\ref{fig:supp:length_vs_metaInfo}).
A similar pattern has been observed in applications of neural embedding to natural language, in which the size of word vectors was found to represent the word's specificity.
Roughly speaking, for word embeddings, the size of a vector indicates a specificity, i.e., the word associated with the vector frequently co-appears with particular context words\autocite{schakel2015measuring}. 
If the word in question is universal, appearing frequently in many different contexts, it would not have a large norm due to a lack of strong association with a particular context.
Likewise, an organization such as Harvard with a small norm appears in many contexts, alongside many different organizations in affiliation trajectories---it is well-connected.
The concavity emerges in part from the relationship between the size of the vector and the expected connectedness of the organization given its size (Fig.~\ref{fig:length}a, inset, $R^{2} = 0.17$).
Large, prestigious, and well-funded research universities such as Princeton and Harvard have smaller vector sizes because they appear in many different contexts compared to more teaching-focused organizations such as NY Medical College, and University of Michigan at Flint.
Some universities, such as the University of Alaska at Fairbanks, have considerably small vectors, which may be a result of its remote location and unique circumstances.

This find a similar convex-curve across many countries such as China, which closely mirrors that of the United States (Fig.~\ref{fig:length}b).
Smaller but scientifically advanced countries such as Australia, and other large populous countries such as Brazil exhibit curves similar to the United States (Fig.~\ref{fig:length}b, inset), though their exact position and size varies.
Other nations exhibit different curves, such as Egypt which has an``S''-shaped curve, meaning that their largest organizations are not as well connected as other countries' largest organizations. 
Similar patterns can be found across many of the 30 of the countries with the most total researchers (Fig.~\ref{fig:concave30}), however we can only speculate on what their variations mean (see \SI).
Neural embeddings encode a relationship between organization size and vector length that is universal across countries but with potentially meaningful nuances, helping the embedding to capture the landscape of global scientific mobility.

%
% Figure - Boomerang
%
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{\figdir/Math/boomerang.pdf}
	\caption{
		\textbf{Size of organization embedding vectors captures prestige and size of organizations.}
		\textbf{a.} Size (L2 norm) of organization embedding vectors compared to the number of researchers for U.S. universities.
		Color indicates the rank of the university from the Times ranking, with 1 being the highest ranked university.
		Uncolored points are universities not listed on a Times ranking.
		A convex-shape emerges, wherein larger universities tend to be more distant from the origin (large L2 norm); however, the more prestigious universities tend to have a small L2 norm.
		The L2 norm is correlated weakly with the ratio of organizations' number of connections (non-zero flows) to the number of researchers ($R^{2} = 0.17$, \textbf{inset}).
		The red line is the line of the best fit, and 99\% confidence intervals are shown, though are too small to be visible.
		\textbf{b.} We find a simliar convex-curve pattern across many countries such as the United States, China, Australia, Brazil, and others (\textit{inset}, and Fig.~\ref{fig:concave30}). 
		Some countries exhibit variants of this pattern, such as Egypt 's ``S''-shaped curve (\textbf{inset}).
		The loess regression lines are shown for each selected country, and for the aggregate of remaining countries, with 99\% confidence intervals.
		Loess lines are also shown for organizations in Australia, Brazil, and Egypt (\textbf{inset}).
	}
	\label{fig:length}
\end{figure}

@@

Neural embeddings offer a novel, data-driven solution for learning an effective and robust functional distance between places based on trajectory data, encoding the complex and multi-faceted nature of mobility.
We demonstrated how the embedding space encodes various aspects of scientific mobility, such as the global and regional geography, shared languages, and the hierarchy of organizational prestige, even without explicit information on these factors.

While scientific mobility is used as a case study, this approach may be broadly applicable to other kinds of mobility data, such as general human migration, transit-network mobility, and more.
Moreover, this approach can be used to learn a functional distance between entities for which no physical analog exists, such as between occupational categories based on individuals' career trajectories.
In addition to providing a functional distance that supports modeling and predicting mobility patterns, the structure of the neural embedding space is amenable to a range of unique applications for studying mobility.
As we have shown, the embedding space allows the visualization of the complex structure of scientific mobility at high resolution across multiple scales, providing a large and detailed map of the landscape of global scientific mobility.
SemAxis~\autocite{an2018semaxis} allows us to quantify abstract notions in mobility such as academic prestige, and can also be generalized to other abstract axes.
Investigation of the structure of the embedding space, such as the vector size, reveals unique and universal patterns based on organization's size and connectedness and could contribute to deeper understandings of scientific mobility.
Neural embedding opens the door for new approaches to represent and study scientific mobility, and also mobility more generally.

In spite of its promise, this approach and analysis has several limitations.
First, the current \textit{skip-gram word2vec} model can not encode directional information, meaning that embedding will be less effective at capturing mobility for which directionality is critical.
Second, the neural embedding approach is most useful in cases of mobility between discrete geographic units such as between countries, cities, and businesses; this approach is less useful in the case of mobility between locations represented using geographic coordinates, such as in the modeling of animal movements.
Neither of these methodological limitations is insurmountable, and future work can aim to incorporate directionality and identify meaningful representations of continuous mobility data.
Finally, the case of scientific mobility presents domain-specific limitations.
Reliance on bibliometric metadata means that we capture only long-term mobility such as migration, rather than the array or more frequent short-term mobility such as conference travel and temporary visits.
The kinds of mobility we do capture---migration and co-affiliation---are treated identically by our model although they are qualitatively different.
Also, our data might further suffer from survivorship bias: researchers at prestigious organizations tend to have more publications, leading to these organizations appearing more frequently in affiliation trajectories.

Mobility and migration are at the core of human nature and history~\autocite{shah2020migration}, driving societal phenomena as diverse as epidemics\autocite{kraemer2020covid, truscott2012epidemicgravity, xia2005measlesgravity} and science and innovation~\autocite{kaiser2018innovation, armano2017innovation, sugimoto2017mostimpact, petersen2018multiscale, franzoni2014advantage, morgan2018prestige, rodrigues2016mobility}.
However, the paradigm of scientific migration may be changing.
Traditional hubs of migration have experienced many politically-motivated policy changes that affect scientific mobility, such as travel restrictions in the U.S. and U.K.~\autocite{chinchilla2018travelban}.
At the same time, other nations, such as China, are growing into major scientific powers and attractors of talent~\autocite{tollefson2018china, cao2020returning}.
Unprecedented health crises such as the COVID-19 pandemic threaten to bring drastic global changes to travel and migration by tightening borders and halting travel.
With the changing paradigm of global mobility we need now, more than ever, new tools and approaches to capture and understand human mobility in order to inform sensible, effective, sustainable, and humane policies.

@@

The performance of the embedding ranking in reconstructing the Times ranking is comparable to that of network-derived measures such as degree strength (Spearman's $\rho = 0.73$, Fig.~\ref{fig:supp:centrality_times_compare}a) and eigencentrality centrality (Spearman's $\rho = 0.76$, Fig.~\ref{fig:supp:centrality_times_compare}b).
However, while both embedding- and network-based measures relate to university prestige, they are qualitatively and quantitatively different.
The embedding-ranking of U.S. universities is less correlated with degree strength (Spearman's $\rho = 0.45$, Fig.~\ref{fig:supp:centrality_semaxis_compare}a) and eigenvector  centrality (Spearman's $\rho = 0.55$) than with the Times ranking itself (Spearman's $\rho = 0.73$, Fig.~\ref{fig:supp:centrality_semaxis_compare}b). 
The embedding ranking over-ranks large research-intensive universities such as North Carolina State University, University of Florida, and Texas A\&M University, whereas the network-derived ranking over-ranks smaller, more specialized universities such as Brandeis University, Yeshiva University, and University of San Francisco. 
This suggests that the embedding encodes information on prestige hierarchy at least as well as a network representation, but that the information encoded is something different from centrality
